---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# govhrcast

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/WB-PIDA-Data-Science-Shop/govhrcast/graph/badge.svg)](https://app.codecov.io/gh/WB-PIDA-Data-Science-Shop/govhrcast)
[![R-CMD-check](https://github.com/WB-PIDA-Data-Science-Shop/govhrcast/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/WB-PIDA-Data-Science-Shop/govhrcast/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Designing a new retirement policy? Planning a hiring freeze? Considering a civil service pay reform? govhrcast is an R package that lets you simulate the workforce and wage bill implications of public sector policy decisions from HRMIS microdata at the contract and personnel level. See [govhr] (https://github.com/WB-PIDA-Data-Science-Shop/govhr) package for the pre-requisite data requirements.

The package provides a modular simulation engine built around a single projection function, simulate_horizon(), which chains four workforce modules — retirement, non-retirement exits, hiring, and promotions/transfers — over successive time periods. Each module is independently configurable, so you can hold some processes at their historical baseline while reforming others.

## Installation

You can install the development version of govhrcast from [GitHub](https://github.com/WB-PIDA-Data-Science-Shop/govhrcast) with:

``` r

# Install the development version from GitHub
remotes::install_github("WB-PIDA-Data-Science-Shop/govhrcast")

```

## Quick start

The package ships with harmonized example data from hypothetical Brazilian state government HRMIS. A minimal status quo projection requires three inputs: a contract module, a personnel module as well as a supplementary allowance module.


```{r}
library(govhrcast)
library(data.table)

# Contract module — one row per contract per snapshot
head(bra_hrmis_contract[ref_date == as.Date("2017-09-01"),
                         .(personnel_id, contract_id, est_id, paygrade,
                           contract_type, gross_salary_lcu)])

# Personnel module — one row per person per snapshot
head(bra_hrmis_personnel[ref_date == as.Date("2017-09-01"),
                          .(personnel_id, age, personnel_tenure,
                            employment_status)])

# Allowance module - one row per person per contract per allowance type per snapshot
head(bra_hrmis_allowance[ref_date == as.Date("2017-09-01"),
                         .(personnel_id, contract_id, allowance_type, allowance_lcu)])


```


## A Status Quo Projection
The status quo run holds every process at its historically observed rate. Retirement follows an age threshold, exits use establishment-level attrition rates estimated from the panel, hiring replicates historical intake rates, and promotions/transfers replicate the observed transition matrix.


```{r}

# Salary scale: mean salary per establishment x paygrade at baseline
salary_scale_dt <- bra_hrmis_contract[
  ref_date == as.Date("2017-09-01"),
  .(gross_salary_lcu = mean(gross_salary_lcu, na.rm = TRUE)),
  by = .(est_id, paygrade)
]

# Historical exit rates — one row per establishment
exit_dt <- estimate_historical_exit_rates(
  panel_contract_dt  = bra_hrmis_contract,
  panel_personnel_dt = bra_hrmis_personnel,
  group_cols         = "est_id"
)

# Historical movement baseline
movement_dt <- govhrcast:::estimate_movement_baseline(
  contract_dt = bra_hrmis_contract,
  group_cols  = c("est_id", "paygrade")
)

statusquo_sim <- simulate_horizon(
  contract_dt     = bra_hrmis_contract,
  personnel_dt    = bra_hrmis_personnel,
  salary_scale_dt = salary_scale_dt,

  retirement_policy = list(
    group_cols   = NULL,
    policy_table = NULL,
    defaults     = list(
      eligibility_type = "age_only",
      min_age          = 60,
      pension_type     = "rate",
      pension_rate     = 0.15,
      ref_wage_col     = "gross_salary_lcu",
      active_types     = c("permanent", "short-term", "fixed-term")
    )
  ),

  exit_policy = list(
    group_cols   = "est_id",
    policy_table = exit_dt,
    defaults     = list(
      exit_rate    = mean(exit_dt$exit_rate, na.rm = TRUE),
      active_types = c("permanent", "short-term", "fixed-term")
    )
  ),

  movement_policy = list(
    group_cols   = c("est_id", "paygrade"),
    policy_table = movement_dt,
    defaults     = list(
      movement_rate      = 0,
      movement_strategy  = "tenure",
      active_types       = "permanent",
      salary_update_rule = "scale"
    )
  ),

  hiring_policy = list(
    mode         = "status_quo",
    group_cols   = c("est_id", "paygrade"),
    salary_scale = salary_scale_dt,
    rate_mult    = 1
  ),

  salary_growth_rate = 0.04,
  ref_date           = as.Date("2017-09-01"),
  age_col            = "age",
  tenure_col         = "personnel_tenure",
  hire_date_col      = "first_employment_date",
  n_periods          = 5,
  salary_col         = "gross_salary_lcu",
  scenario_name      = "status_quo"
)

```


### Projection output
simulate_horizon() returns a horizon object. The $summary_dt element contains one row per simulation period with key workforce and wage bill statistics.


```{r}

statusquo_sim$summary_dt

```


### How it works

simulate_horizon() chains four options modules (set any to NULL if you prefer) over each projection period in a fixed sequence:

```{r workflow-diagram, echo = FALSE, results = "asis"}
cat("
\`\`\`mermaid
flowchart TD
    A[\"Contract & Personnel data\"] --> B[\"simulate_retirement()\"]
    B --> C[\"simulate_exits()\"]
    C --> D[\"simulate_hiring()\"]
    D --> E[\"simulate_promotions_transfers()\"]
    E --> F{Next period?}
    F -- Yes --> B
    F -- No --> G[\"horizon object\"]
\`\`\`
")
```


Each module is independently configurable through a policy list and leaves the others unchanged, so you can reform one dimension of the workforce system while holding everything else at its baseline.

| Module | What it models |
| --- | --- |
| simulate_retirement() | Age- and/or tenure-based retirement eligibility; defined benefit, defined contribution, rate, flat, and hybrid pension formulas |
| simulate_exits() | Non-retirement attrition at scalar or group-varying rates; random or rank-based exit selection |
| simulate_hiring() | Demand-driven hiring under four modes: status quo, flow (retirement replacement), stock (headcount target), and combined |
| simulate_promotions_transfers() | Vertical promotions and lateral transfers via empirical transition matrices or user-specified rates; four mover selection strategies |


## What can you model: 

`govhrcast` is designed around the kinds of questions that arise in public sector fiscal and workforce reform work:

- **Fiscal consolidation** — what happens to the wage bill if we freeze hiring for three years and rely on natural attrition to reduce headcount?
- **Retirement waves** — how many workers will become eligible for retirement over the next decade, and what is the pension liability?
- **Replacement policy** — if we replace only 80% of retirees, how long until headcount stabilises, and at what wage bill level?
- **Pay reform** — how does a salary scale restructuring interact with planned hiring and expected attrition over five years?
- **Promotion acceleration** — what is the wage bill cost of doubling promotion rates into senior grades as part of a civil service reform?

Scenarios are compared by running `simulate_horizon()` multiple times with different policy lists and inspecting the `$comparison` element of the returned `horizon` object.



## Policy specification
Every module follows the same three-slot policy format:


```{r}

list(
  group_cols   = "est_id",          # grouping variables — NULL for uniform policy
  policy_table = exit_dt,           # group-specific overrides — NULL for defaults only
  defaults     = list(              # fallback values for all groups
    exit_rate    = 0.05,
    active_types = c("permanent", "short-term", "fixed-term")
  )
)

```

Groups absent from policy_table automatically inherit the corresponding value from defaults. This means you can calibrate policy for the establishments you have data for and fall back gracefully for the rest.


## Documentation
Full function-level documentation is available via ?simulate_horizon, ?simulate_retirement, ?simulate_exits, ?simulate_hiring, and ?simulate_promotions_transfers.

Vignettes covering status quo projections, reform scenario comparisons, and pension liability estimation are in development.

## License
MIT © Ifeanyi Edochie, Galileu Kim
